Efficient Handling of String-Number Conversion

### 2019-11-22 evaluation.tex
minor update to evalutaion
Update table; Revised evaluation.

\section{Evaluation}
\label{section:evaluation}

We have implemented our string constraint solving procedure in a tool called {\tool}. {\tool} is implemented on top of the SMT solver Z3~\cite{z3} as a theory solver. In this way we can concentrate on solving constraints connected by conjunctions and let Z3 to handle other boolean connectives. It also enables us to solve not only formulae over string constraints, but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

@@

We compare {\tool} with other state-of-the-art string solvers, namely CVC4~(relese-1.7)~\cite{cvc4Tool}, Z3~(\texttt{dde8da853})~\cite{z3}, \textsf{Z3-str3}~(\texttt{1a9dfc5e8})~\cite{zheng2013z3}, and \textsf{Trau+}~\cite{abdulla2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints, which occurs in most of our benchmarks. We also do not compare with ABC~\cite{aydin2018parameterized} (a model counter for string constraints) and Ostrich~\cite{chen2019decision}, because most of our benchmarks contain non-trivial equality constraints, which is not fully supported by those solvers.

We perform two sets of experiments. In the first set of experiment, we compare with other tools on benchmarks over basic string constraints. We use the following benchmark examples in this experiments:
\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.
	\item APLAS~\cite{aplas} includes 600 hand-drafted tests consisting of only equality and integer length constraints.
	\item LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website.
	\item StringFuzz~\cite{blotsky2018stringfuzz}  is generated from the fuzz testing tool of the same name.
	\item $\text{cvc4}_{\text{pred}}$ and $\text{cvc4}_{\text{term}}$ are obtained from the CVC4 group~\cite{termEQ}. These benchmarks contain a small amount of string-number conversion constraints (< $5\%$).
\end{itemize}

@@

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3/Z3str3 on Basic String Constraint benchmarks.}
\scalebox{0.85}{
\begin{tabular}{|l r | r r r r r|}
\hline
\multicolumn{2}{|c}{}                  & {\tool} & CVC4  &Z3 & Z3-str3 & Trau+  \\ \hline
\multirow{3}{*}{PyEx}		& SAT      &  21205  & 19763 & 18490 &  3514 & 19149 \\ 
							& UNSAT    &   3963  &  3834 &  3855 &  4521 & 3828 \\
							& $\times$ &    253  &  1824 &  3076 & 17386 & 2444 \\ \hline
\multirow{3}{*}{APLAS}		& SAT      &    135  &   205 &    19 &    52 & 132 \\
							& UNSAT    &    343  &   221 &   100 &   118 & 82 \\
							& $\times$ &    122  &   174 &   481 &   430 & 386 \\ \hline
\multirow{3}{*}{LeetCode}	& SAT      &    842  &   859 &   881 &   651 & 778 \\
							& UNSAT    &   1824  &  1785 &  1785 &  1787 & 1827 \\
							& $\times$ &      0  &    22 &     0 &   228 & 61 \\ \hline
\multirow{3}{*}{StringFuzz}	& SAT      &    528  &   671 &   316 &   605 & 688 \\
							& UNSAT    &    320  &   240 &   206 &   217 & 339 \\
							& $\times$ &    217  &   154 &   543 &   243 & 38 \\\hline
\multirow{3}{*}{cvc4\textsubscript{pred}} & SAT & 13 & 11 &   12 &    10 & 247 \\
							& UNSAT    &    820  &   818 &   807 &   778 & 500 \\
							& $\times$ &      2  &     6 &    16 &    47 & 88 \\ \hline
\multirow{3}{*}{cvc4\textsubscript{term}} & SAT & 13 & 8 &     6 &    16 & 311 \\
							& UNSAT    &   1031  &   936 &  1020 &   962 & 596 \\
							& $\times$ &      1  &   101 &    19 &    67 & 138 \\ \hline
\multirow{3}{*}{Total} 		& SAT      &  22736  &  21517 & 19724 &  4848 & 21305  \\
							& UNSAT    &   8301  &   7834 &  7773 &  8383 & 7172  \\
							& $\times$ &    595  &   2281 &   4135 & 18401 & 3155  \\\hline	
\end{tabular}}
\label{table:base_benchmark}
\end{table}

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, and Z3/Z3str3 on Basic String Constraints + String-Number Conversion benchmark.}
\scalebox{0.8}{
\begin{tabular}{|l r | r r r r|}
\hline
\multicolumn{2}{|c}{}                   			& {\tool} & CVC4   &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{String-Number-Leetcode}	 & SAT      &   2413  &  1543 &  1025  &   333 \\ 
										 & UNSAT    &  16511  & 15676 & 17234  & 15543 \\
										 & $\times$ &      3  &  1708 &   668  &  3051 \\ \hline 
\multirow{3}{*}{String-Number-PythonLib} & SAT      &    424  &   408 &  1802  &   389 \\ 
										 & UNSAT    &   2056  &   660 &   744  &   729 \\
										 & $\times$ &    166  &  1578 &   100  &  1528 \\ \hline
\multirow{3}{*}{String-Number-JavaScript}& SAT      &     20  &     6 &    16  &     4 \\ 
										 & UNSAT    &      0  &     0 &     0  &     0 \\
										 & $\times$ &      0  &    14 &     4  &    16 \\ \hline \hline
\multirow{3}{*}{Total}					 & SAT      &   2857  &  1957 &  2843  &   726 \\ 
										 & UNSAT    &  18567  & 16336 & 17978  & 16272 \\
										 & $\times$ &    169  &  3300 &   772  &  4595 \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}

From Table~\ref{table:base_benchmark}, we can see that the performance of {\tool} is as good as the most competitive tools such as CVC4 and Z3 on basic string constraints. In all of the benchmarks, {\tool} ranked either the 1st or the 2nd on the number of solved (SAT+UNSAT) cases. If we look into scrutiny, for the APLAS and StringFuzz benchmarks, there is a significant distance between {\tool} and the top-ranked tool on the SAT cases. But we believe this is not very important because both these two benchmarks are artificial, either hand-crafted or randomly produced from a fuzzer. The more critical ones are those come from program analysis, which {\tool} has comparable performance to the best tool.

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The total number of failed cases is only one-fourth to Z3, which is ranked the 2nd. In fact, most of the failed tests comes from the analysis of one Python core library function that converts an IPv6 address to a number. We generated 2028 tests from this function, and among them around 150 cases that are too difficult for all solvers.  If we exclude results on this function, then {\tool} has only $185$ failed case in total. This is significantly better than the 2nd place tool Z3, which fails $893$ cases in total. \yfc{This paragraph needs to be revised when new data of Z3-Trau is available. Note that z3seq is much more better now. Also is z3-trau.}

\hide{ % data for helping to write comparison
benchmark_path: lib_int-ipaddress__ip_int_from_string
benchmark_size: 2028
(tool-check_date, #sat, #unsat, #timeout, #unknown, #exception/error, #misc)
(cvc4-20191120, 96, 367, 1565, 0, 0, 0)
(z3-trau-20191122, 135, 1739, 154, 0, 0, 0)
(z3seq-20191118, 1509, 431, 88, 0, 0, 0)
(z3str3-20191118, 191, 440, 1330, 5, 0, 62)

benchmark_path: leetcode_int-
benchmark_size: 18927
(tool-check_date, #sat, #unsat, #timeout, #unknown, #exception/error, #misc)
(z3seq-20191118, 1025, 17234, 668, 0, 0, 0)
(cvc4-20191120, 1543, 15676, 1708, 0, 0, 0)
(cvc4-20191118, 1766, 15612, 1549, 0, 0, 0)
(z3-trau-20191122, 2413, 16511, 3, 0, 0, 0)
(z3str3-20191118, 333, 15543, 2031, 952, 0, 68)

benchmark_path: lib_int-
benchmark_size: 2646
(tool-check_date, #sat, #unsat, #timeout, #unknown, #exception/error, #misc)
(cvc4-20191118, 586, 666, 1394, 0, 0, 0)
(z3seq-20191118, 1802, 744, 100, 0, 0, 0)
(cvc4-20191120, 408, 660, 1578, 0, 0, 0)
(z3str3-20191118, 389, 729, 1392, 70, 0, 66)
(z3-trau-20191122, 424, 2056, 166, 0, 0, 0)
}

\begin{table}[h]
\centering
\caption{Comparison of {\tool}, CVC4, and Z3/Z3str3 with checkLuhn problems of 2 to 12 loops.}
\scalebox{0.9}{
\begin{tabular}{| c | c c c c|}
\hline
\# of Loops & {\tool} 	  &  CVC4    	 &       Z3   & Z3-str3 \\ \hline
2 			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(0.66s) & \textbf{SAT}(0.2s)  & $\times$ \\
3  			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(1.78s) & \textbf{SAT}(0.1s)  & $\times$ \\
4 			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(6.96s) &   $\times$  		& $\times$ \\
5 			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(17.2s) &   $\times$  		& $\times$ \\
6 			& \textbf{SAT}(0.16s)  &  $\times$  		  & \textbf{SAT}(0.13s) & $\times$ \\
7 			& \textbf{SAT}(0.24s)  &  $\times$    		  &   $\times$  		& $\times$ \\
8 			& \textbf{SAT}(0.31s)  &  $\times$   		  & \textbf{SAT}(0.37s) & $\times$ \\
9 			& \textbf{SAT}(0.26s)  &  $\times$    		  &   $\times$  		& $\times$ \\
10 			& \textbf{SAT}(0.21s)  &  $\times$    		  & \textbf{SAT}(0.53s) & $\times$ \\
11 			& \textbf{SAT}(0.23s)  &  $\times$    		  & \textbf{SAT}(0.37s) & $\times$ \\
12 			& \textbf{SAT}(0.33s)  &  $\times$     		  & \textbf{SAT}(0.40s) & $\times$ \\ \hline
\end{tabular}}
\label{table:checkLuhn}
\end{table}

>>>>

\section{Implementation and Evaluation}
\label{section:evaluation}

We have implemented our string constraint solving procedure in a tool called {\tool}. {\tool} is implemented as a theory solver of the SMT solver Z3~\cite{z3} . In this way, we can concentrate on solving constraints connected by conjunctions and let Z3 handle other boolean connectives. It also enables us to solve not only formulae over string constraints but also the combination of string constraints with constraints over other theories that Z3 supports. Furthermore, this approach allows us to more effectively handle the arithmetic constraints that are generated by the under-approximation module and eliminate the need to have our own parser for input formulae.

@@

We compare {\tool} with other state-of-the-art string solvers, namely CVC4~(45bcf28a)~\cite{cvc4Tool}, Z3~(\texttt{d95b549ff})~\cite{z3}, \textsf{Z3-str3}~(\texttt{d95b549ff})~\cite{zheng2013z3}, and \textsf{Trau+}~\cite{abdulla2019chain}. We do not compare with Sloth \cite{sloth} since it does not support length constraints, which occurs in most of our benchmarks. We also do not compare with ABC~\cite{aydin2018parameterized} (a model counter for string constraints) and Ostrich~\cite{chen2019decision}, because most of our benchmarks contain non-trivial equality constraints, which is not fully supported by those solvers.

We perform two sets of experiments. In the first set of experiments, we compare with other tools on benchmarks over basic string constraints. We use the following benchmark examples in these experiments:
\begin{itemize}
	\item PyEx~\cite{pyex} is from running the symbolic executor PyEx over some Python packages.
	\item APLAS~\cite{aplas} includes 600 hand-drafted tests consisting of only equality and integer length constraints.
	\item LeetCode is from running PyEx over some sample code collected from the LeetCode~\cite{LeetCode} website.
	\item StringFuzz~\cite{blotsky2018stringfuzz}  is generated from the fuzz testing tool of the same name.
	\item $\text{cvc4}_{\text{pred}}$ and $\text{cvc4}_{\text{term}}$ are obtained from the CVC4 group~\cite{termEQ}. These benchmarks contain a small amount of string-number conversion constraints (< $5\%$).
\end{itemize}

@@

\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, Z3, Z3-str3 and Trau+ on Basic String Constraint benchmarks.}
\scalebox{0.85}{
\begin{tabular}{|l r | r r r r r|}
\hline
\multicolumn{2}{|c}{}                  & {\tool} & CVC4  &Z3 & Z3-str3 & Trau+  \\ \hline
\multirow{3}{*}{PyEx}		& SAT      &  21205  & 19763 & 16528 &  3030 & 19149 \\ 
							& UNSAT    &   3963  &  3834 &  3831 &  3836 & 3828 \\
							& $\times$ &    253  &  1824 &  5062 & 18555 & 2444 \\ \hline
\multirow{3}{*}{APLAS}		& SAT      &    135  &   205 &    12 &    37 & 132 \\
							& UNSAT    &    343  &   221 &   100 &   111 & 82 \\
							& $\times$ &    122  &   174 &   488 &   452 & 386 \\ \hline
\multirow{3}{*}{LeetCode}	& SAT      &    842  &   859 &   881 &   670 & 778 \\
							& UNSAT    &   1824  &  1785 &  1785 &  1780 & 1827 \\
							& $\times$ &      0  &    22 &     0 &   216 & 61 \\ \hline
\multirow{3}{*}{StringFuzz}	& SAT      &    528  &   671 &   264 &   491 & 688 \\
							& UNSAT    &    320  &   240 &   186 &   192 & 339 \\
							& $\times$ &    217  &   154 &   615 &   382 & 38 \\\hline
\multirow{3}{*}{cvc4\textsubscript{pred}} & SAT & 13 & 11 &   12 &     8 & 247 \\
							& UNSAT    &    820  &   818 &   808 &   772 & 500 \\
							& $\times$ &      2  &     6 &    15 &    55 & 88 \\ \hline
\multirow{3}{*}{cvc4\textsubscript{term}} & SAT & 13 & 8 &     5 &    17 & 311 \\
							& UNSAT    &   1031  &   936 &  1021 &   958 & 596 \\
							& $\times$ &      1  &   101 &    19 &    70 & 138 \\ \hline
\multirow{3}{*}{Total} 		& SAT      &  22736  & 21517 & 17702 &  4253 & 21305  \\
							& UNSAT    &   8301  &  7834 &  7731 &  7649 & 7172  \\
							& $\times$ &    595  &  2281 &  6199 & 19730 & 3155  \\\hline	
\end{tabular}}
\label{table:base_benchmark}
\end{table}

\hide{
% wider table for 30s timeout data
\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, Z3, and Z3-str3 on String-Number Conversion benchmark.}
\scalebox{0.8}{
\begin{tabular}{|l r | r r r r|| r r r r|}
\hline
\multicolumn{2}{|c}{}          			   & {\tool} & CVC4  &    Z3  & Z3-str3 & {\tool} & CVC4  &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{Leetcode}		& SAT      &   2413  &  1543 &  1892  &   217 \\ 
								& UNSAT    &  16511  & 15676 & 16105  & 15374 \\
								& $\times$ &      3  &  1708 &   930  &  3336 \\ \hline 
\multirow{3}{*}{PythonLib}		& SAT      &    424  &   408 &   797  &   201 \\ 
								& UNSAT    &   2056  &   660 &   724  &   644 \\
								& $\times$ &    166  &  1578 &  1125  &  1801 \\ \hline
\multirow{3}{*}{JavaScript}		& SAT      &     20  &     6 &    16  &     4 \\ 
								& UNSAT    &      0  &     0 &     0  &     0 \\
								& $\times$ &      0  &    14 &     4  &    16 \\ \hline \hline
\multirow{3}{*}{Total}			& SAT      &   2857  &  1957 &  2705  &   422 \\ 
								& UNSAT    &  18567  & 16336 & 16829  & 16018 \\
								& $\times$ &    169  &  3300 &  2059  &  5153 \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}
}

% latest backup of table 2
\begin{table}[h]
\centering
\caption{Results of {\tool}, CVC4, Z3, and Z3-str3 on String-Number Conversion benchmark.}
\scalebox{0.8}{
\begin{tabular}{|l r | r r r r|}
\hline
\multicolumn{2}{|c}{}          			   & {\tool} & CVC4  &    Z3  & Z3-str3 \\ \hline
\multirow{3}{*}{Leetcode}		& SAT      &   2413  &  1543 &  1892  &   217 \\ 
								& UNSAT    &  16511  & 15676 & 16105  & 15374 \\
								& $\times$ &      3  &  1708 &   930  &  3336 \\ \hline 
\multirow{3}{*}{PythonLib}		& SAT      &    424  &   408 &   797  &   201 \\ 
								& UNSAT    &   2056  &   660 &   724  &   644 \\
								& $\times$ &    166  &  1578 &  1125  &  1801 \\ \hline
\multirow{3}{*}{JavaScript}		& SAT      &     20  &     6 &    16  &     4 \\ 
								& UNSAT    &      0  &     0 &     0  &     0 \\
								& $\times$ &      0  &    14 &     4  &    16 \\ \hline \hline
\multirow{3}{*}{Total}			& SAT      &   2857  &  1957 &  2705  &   422 \\ 
								& UNSAT    &  18567  & 16336 & 16829  & 16018 \\
								& $\times$ &    169  &  3300 &  2059  &  5153 \\ \hline
\end{tabular}}
\label{table:str_int_benchmark}
\end{table}

From Table~\ref{table:base_benchmark}, we can see that the performance of {\tool} is as good as the most competitive tools such as CVC4 and Z3 on basic string constraints. In all of the benchmarks, {\tool} ranked either the 1st or the 2nd on the number of solved (SAT+UNSAT) cases. If we look into scrutiny, for the APLAS and StringFuzz benchmarks, there is a significant distance between {\tool} and the top-ranked tool on the SAT cases. But we believe this is not very important because both these two benchmarks are artificial, either hand-crafted or randomly generated. The more critical ones are those come from program analysis, in which {\tool} has comparable performance to the best tool.

From Table~\ref{table:str_int_benchmark}, we can see that {\tool} significantly outperforms all other tools. The total number of failed cases is only a half to Z3, which is ranked the 2nd. In fact, most of the failed tests come from the analysis of one Python core library function that converts an IPv6 address to a number. We generated 2028 tests from this function, and among them around 1000 cases that are too difficult for all solvers.  If we exclude results on this function, then {\tool} has only $185$ failed cases in total. This is significantly better than the 2nd place tool Z3, which fails $893$ cases in total. \todo{Check and revise the numbers after table updated.}

\todo{Mention about (1) checkLuhn cases (2) out tool solves much more cases if timeout is set to 30s. (3) update table with data of timeout=30s}

\begin{table}[h]
\centering
\caption{Comparison of {\tool}, CVC4, and Z3/Z3str3 with checkLuhn problems of 2 to 12 loops.}
\scalebox{0.9}{
\begin{tabular}{| c | c c c c|}
\hline
\# of Loops & {\tool} 			   &  CVC4    	 		  &       Z3   			& Z3-str3 \\ \hline
2 			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(0.66s) & \textbf{SAT}(0.23s) & $\times$ \\
3  			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(1.78s) & \textbf{SAT}(0.13s) & $\times$ \\
4 			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(6.96s) &   $\times$  		& $\times$ \\
5 			& \textbf{SAT}(<0.1s)  &  \textbf{SAT}(17.2s) &   $\times$  		& $\times$ \\
6 			& \textbf{SAT}(0.16s)  &  $\times$  		  & \textbf{SAT}(0.14s) & $\times$ \\
7 			& \textbf{SAT}(0.24s)  &  $\times$    		  &   $\times$  		& $\times$ \\
8 			& \textbf{SAT}(0.31s)  &  $\times$   		  & \textbf{SAT}(0.37s) & $\times$ \\
9 			& \textbf{SAT}(0.26s)  &  $\times$    		  &   $\times$  		& $\times$ \\
10 			& \textbf{SAT}(0.21s)  &  $\times$    		  & \textbf{SAT}(0.62s) & $\times$ \\
11 			& \textbf{SAT}(0.23s)  &  $\times$    		  & \textbf{SAT}(0.47s) & $\times$ \\
12 			& \textbf{SAT}(0.33s)  &  $\times$     		  & \textbf{SAT}(0.47s) & $\times$ \\ \hline
\end{tabular}}
\label{table:checkLuhn}
\end{table}
